---
format: 
        gfm:
                message: false
                warning: false
                fig-dpi: 300
                fig-height: 6
                fig-width: 8
editor: source
---

# sentiment analysis for star wars


## data

Load packages, functions, and data for analyzing sentiment in the scripts of the original Star Wars trilogy

```{r}
#| label: setup

# load packages
library(tidyverse)
library(tidytext)
library(ggthemes)
library(textdata)
library(sentimentr)
library(ggrepel)
library(ggforce)
library(plotly)

# load and set ggplot theme
source("src/functions.R")

theme_set(
        theme_phil()+
                theme(
                        strip.text.x = element_text(size = 12)
                )
)

# tokenize and remove stopwords
load_stopwords()

# read in starwars text
starwars =
        load_data() |>
        clean_data()


```


## sentiment

Calculate sentiment via **sentimentr**.

```{r}

# use sentimentr
starwars_sentiment = 
        starwars |>
        get_sentences()  |>
        add_sentiment()

```

# sentiment by dialogue

what pieces of dialogue in star wars have the highest average positive/negative sentiment?

## positive

```{r}

starwars_sentiment |>
        group_by(document, index = line_number %/% 4) |>
        summarize(sentiment = mean(sentiment),
                  characters = paste(unique(character), collapse = '\n'),
                  dialogue = paste(dialogue, collapse = '\n'),
                  .groups = 'drop') |>
        select(document, characters, dialogue, sentiment) |>
        slice_max(sentiment, n = 10) |>
        mutate_if(is.numeric, round, 2) |>
        gt::gt()

```


## negative

```{r}

starwars_sentiment |>
        group_by(document, index = line_number %/% 4) |>
        summarize(sentiment = mean(sentiment),
                  characters = paste(unique(character), collapse = '\n'),
                  dialogue = paste(dialogue, collapse = '\n'),
                  .groups = 'drop') |>
        select(document, characters, dialogue, sentiment) |>
        slice_min(sentiment, n = 10) |>
        mutate_if(is.numeric, round, 2) |>
        gt::gt()

```


# sentiment by character

Plot average sentiment across main speaking roles in the original trilogy.

```{r}
#| fig-height: 8
#| fig-width: 8

starwars |>
        plot_avg_sentiment(
                top_n = 20,
                scales = "free_y"
        )
```


# sentiment by line

Plot running sentiment for a character across the script.

```{r}

# plot sentiment for all characters
starwars_sentiment |>
        filter(document == 'a new hope') |>
        filter(character == 'LUKE') |>
        plot_character_lines_with_sentiment()

```


```{r}

# plot sentiment for all characters
starwars_sentiment |>
        filter(document == 'a new hope') |>
        filter(character == 'HAN') |>
        plot_character_lines_with_sentiment()

```


```{r}

# plot sentiment for all characters
starwars_sentiment |>
        filter(document == 'a new hope') |>
        filter(character == 'LEIA') |>
        plot_character_lines_with_sentiment()

```

